---
title: "Lab 4"
author: "Ashley Anderson & Hexiang Zhang"
date: "10/24/2019"
output: html_document
---

Owner Username: ashleyna7  
Partner Username: hexzha  
Repo Name: ds202_lab4  


## Loading Data: Cyclone Football 

```{r}
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)
```

```{r,echo=FALSE,results='hide'}
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
str(offense)
bio <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
str(bio)
```

## Part one: Cleaning data
The three Excel sheets are loaded into R but they are not useful to us before some cleaning. Manipulate the three imported data frames, so that the cleaned data satisfy the following:

1. The player names and opponent names are factors.

```{r}
defClean <- defense
offClean <- offense
bioClean <- bio

defClean$Name <- as.factor(defClean$Name)
offClean$Name <- as.factor(offClean$Name)
bioClean$Name <- as.factor(bioClean$Name)

defClean$Opponent_Opponent <- as.factor(defClean$Opponent_Opponent)
offClean$Opponent_Opponent <- as.factor(offClean$Opponent_Opponent)
```

2. All offensive and defensive statistics are numerical. The `Weight` column in Biography is also numeric. (Hint: look at `?mutate_at` if you want to save writing some code). 

```{r}
defense_names <- c(colnames(defClean[,3:11]))
defClean[defense_names] <- sapply(defClean[defense_names],as.numeric)

offense_names <- c(colnames(offClean[,3:12]))
offClean[offense_names] <- sapply(offClean[offense_names],as.numeric)

bioClean$Weight <- as.numeric(bioClean$Weight)
```

3. Separate the `Height` column into two columns. Make sure the newly created columns are numeric. Then create one column that represents `Height` (Hint: Use a unit that works well). Drop any irrelavant columns from this process. 

```{r}
library(tidyverse)
bioClean <- bio %>%
  separate(Height, c("Feet", "Inches")) %>%
  mutate_each_(funs(as.numeric), c("Feet", "Inches")) %>%
  group_by(Feet, Inches) %>%
  mutate(Height=(Feet*12)+Inches)

head(bioClean$Height)
```

4. By cleaning the three data frames in R, at the end of this step you should have the following three clean data frames: `defClean`, `offClean`, and `bioClean`. Print the structures of these data frames. **Make sure your data types make sense for each variable.**

```{r}
str(defClean)
str(offClean)
str(bioClean)
```


## Part two: Tidying (A)
For the following questions, work on the cleaned data frames `defClean`, `offClean`, and `bioClean`.

1. Reformat the `defClean` data frame into a tidy format using `gather`. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a **new key column named `Statistic`.** 

```{r}
defClean <- defClean %>%
  gather(key=Statistic, value=Value, 3:11)
```

2. Create a histogram for each defensive statistic (make sure to explore different binwidths). Use `facet_wrap` to draw all histograms in a single plot. What defensive skills are rare?

```{r}
ggplot(defClean, aes(Value, fill=Statistic)) + geom_histogram(breaks=seq(0,6,by=1)) + facet_wrap(~Statistic) + labs(title="Histogram of ISU Football Defensive Statistics", x="Value", y="Count")
```
We can see that Turnover_FF, Turnover_FR, and Turnover_INT are all rare as the graph does not show any other bins than than 0 bin, indicating these defensive skills rarely happen. 


3. Did ISU have better defense against West Virginia or Kansas State? Answer this question by creating a scatterplot with x-axis being the number of solo tackles (by each player) against West Virginia, and y-axis being that against Kansas State. A large number of solo tackles indicates good defense. 

      (**Hint**: From step 1 we have gathered our `defClean` into 4 columns. \
       Now we only want to focus on solo tackles (hint...`filter`). 
      Look at your data structure. What form do we need our data to have in order to easily answer this question? (hint...`spread`)). 
      
```{r}
defClean %>%
  filter(Statistic=="Tackles_Solo") %>% 
  spread(key=Opponent_Opponent, value=Value) %>%
  rename(KansasState = "Kansas State", WestVirginia = "West Virginia") %>%
  drop_na(KansasState, WestVirginia) %>%
  ggplot(aes(x=WestVirginia, y=KansasState)) + geom_point() + labs(title="Defensive Tackles: West Virginia vs. Kansas State", x="West Virginia", y="Kansas State") + xlim(0,9) + ylim(0,9)

```
From the plot, we can say that ISU has better defense against Kansas State than West Virginia. 

## Part two: Tidying  (B)
Now let's turn our attention to the player biography `bioClean`.

1. Separate the `Hometown` column into two columns, namely the city and the state. Print the first few rows of these two columns. (Hint: look at the `sep=` argument of `separate`)

```{r}

```

2. How many players are there from each state?

```{r}

```


## Part three: Joining data frames

1. Find out where the offensive players are from by joining the offensive data and player biography data appropriately.

```{r}

```

2. Is there a relationship between the weight of a player and the receiving yards (`Receiving_YDS`)?

```{r}

```

3. Explore more. Raise a question that can be answered only by looking at two or more data frames. Then answer that question.

```{r}

```

